iSEE: Interactive SummarizedExperiment Explorer

### 2018-04-27 manuscript/main.tex
further flesh with some more info, plus cleaning up template leftovers

%\section*{Demonstration data sets and tours}
\section*{Use cases} % FM: used often in other software articles in F1000res

To demonstrate the functionality and flexibility of the iSEE package, we applied our method to a number of publicly available datasets.
Details are provided in the tour subfolder of the GitHub repository for this manuscript (https://github.com/LTLA/iSEE2018), and for each dataset a combination of _app.R and _tour.txt is provided.
An instance of iSEE running on the respective datasets is provided as a stand-alone webpages, for demonstration purposes without the need of software installation.

\subsection*{Single-cell plate-based RNA-seq}
% TODO: better title, or do we drop this allen section altogether?

We demonstrate the application of iSEE to examine a subset of 379 cells from of mouse visual cortex \citep{tasic2016allen}.

@@

\subsection*{Single-cell droplet-based RNA-seq}
% TODO: better title, or do we drop the allen section altogether (above)?

We also demonstrate the use of iSEE to explore a larger data set of 4,000 PBMC single-cell RNA sequencing dataset from 10X Genomics \citep{zheng2017massively}.

% embed the allen demo data set
http://shiny.imbei.uni-mainz.de:3838/iSEE_PBMC4k/

\subsection*{Bulk RNA-seq from TCGA}

We also demonstrate how iSEE seamlessly scales up with larger data sets as the TCGA RNA-seq collection of measurements for 23,368 genes across 7,706 tumor samples \citep{piccolo2015TCGA}.

% TODO: embed the TCGA demo data set
http://shiny.imbei.uni-mainz.de:3838/iSEE_TCGA/

\subsection*{CyTOF}

Lastly, we also showcase the use of iSEE to vizualise and refine gating a gating analysis, followed by the investigation of a functional marker between two sample groups, using a published mass cytometry (CyTOF) data set \citep{bodenmiller2012cytof}.

% TODO: embed the TCGA demo data set
http://shiny.imbei.uni-mainz.de:3838/iSEE_cytof/

\section*{Conclusion}
iSEE provides a general interactive interface for visual exploration of a diverse range of `-omics datasets.
We provide a number of demonstrations at https://github.com/LTLA/iSEE2018, using iSEE to explore the TCGA RNA-seq collection \citep{piccolo2015TCGA}, the 4,000 PBMC single-cell RNA sequencing dataset from 10X Genomics \citep{zheng2017massively} and a published mass cytometry dataset consisting of over 170,000 cells \citep{bodenmiller2012cytof}.
% TODO: repurpose the conclusion if we showcase the examples in earlier sections
% TODO: sell the idea that for every dataset it is possible to explore and showcase feats of interest

@@

\subsection*{Data and software availability}
- source code
- archived source
- package at...
- shiny at...
- docker/singularity containers
- conda pkg
- Software license: MIT

@@

\bigskip
References can be listed in any standard referencing style that uses a numbering system
(i.e. not Harvard referencing style), and should be consistent between references within
a given article.

>>>>

%\section*{Demonstration data sets and tours}
\section*{Use cases} % FM: used often in other software articles in F1000res
To demonstrate the functionality and flexibility of the iSEE package, we applied our method to a number of publicly available datasets.
Details are provided in the tour subfolder of the GitHub repository for this manuscript (https://github.com/LTLA/iSEE2018), and for each dataset a combination of _app.R and _tour.txt is provided.
An instance of iSEE running on the respective datasets is provided as a stand-alone webpages, for demonstration purposes without the need of software installation.

\subsection*{Single-cell plate-based RNA-seq}
% TODO: better title, or do we drop this allen section altogether?
We demonstrate the application of iSEE to examine a subset of 379 cells from of mouse visual cortex \citep{tasic2016allen}.
This example can be reproduced also by calling

library(iSEE)
example(iSEE, ask = F)

@@

\subsection*{Single-cell droplet-based RNA-seq}
% TODO: better title, or do we drop the allen section altogether (above)?
We also demonstrate the use of iSEE to explore a larger data set of 4,000 PBMC single-cell RNA sequencing dataset from 10X Genomics \citep{zheng2017massively}.
% TODO: take out the juice from what the tour does, and maybe one figure/screenshot in each case?
% embed the allen demo data set
http://shiny.imbei.uni-mainz.de:3838/iSEE_PBMC4k/

\subsection*{Bulk RNA-seq from TCGA}
We also demonstrate how iSEE seamlessly scales up with larger data sets as the TCGA RNA-seq collection of measurements for 23,368 genes across 7,706 tumor samples \citep{piccolo2015TCGA}.
% TODO: take out the juice from what the tour does, and maybe one figure/screenshot in each case?
% TODO: embed the TCGA demo data set
http://shiny.imbei.uni-mainz.de:3838/iSEE_TCGA/

\subsection*{CyTOF}
Lastly, we also showcase the use of iSEE to vizualise and refine gating a gating analysis, followed by the investigation of a functional marker between two sample groups, using a published mass cytometry (CyTOF) data set \citep{bodenmiller2012cytof}.
% TODO: take out the juice from what the tour does, and maybe one figure/screenshot in each case?

http://shiny.imbei.uni-mainz.de:3838/iSEE_cytof/

\section*{Conclusion}
iSEE provides a general interactive interface for visual exploration of a diverse range of `-omics datasets.
We provide a number of demonstrations at https://github.com/LTLA/iSEE2018, using iSEE to explore the TCGA RNA-seq collection \citep{piccolo2015TCGA}, the 4,000 PBMC single-cell RNA sequencing dataset from 10X Genomics \citep{zheng2017massively} and a published mass cytometry dataset consisting of over 170,000 cells \citep{bodenmiller2012cytof}.
% TODO: repurpose the conclusion if we showcase the examples in earlier sections
% TODO: sell the idea that for every dataset it is possible to explore and showcase feats of interest. Something like...
We expect the functionality and flexibility of the iSEE package to be essential when exploring and showcasing relevant features of any dataset which can be represented in the form of a SummarizedExperiment object, in such a way that a public instance of iSEE can ideally accompany any publication, e.g. reproducing the manuscript figures, but also allowing any user to foster hypothesis generation with efficient interactive and reproducible data analysis.

@@

\subsection*{Data and software availability}
The iSEE package is available at https://bioconductor.org/packages/iSEE

Source code of the development version is available at https://github.com/csoneson/iSEE

Archived source code as at the time of publication is available at % TODO

An instance of the package running on a public Shiny Server is available at http://shiny.imbei.uni-mainz.de:3838/iSEE/

% TODO: docker/singularity containers

% TODO:conda pkg
% The iSEE package is also distributed via the bioconda channel for the conda package manager at...https://anaconda.org/bioconda/bioconductor-isee % still needs to be finalized!!

Software license: MIT

@@

\bigskip
% References can be listed in any standard referencing style that uses a numbering system
% (i.e. not Harvard referencing style), and should be consistent between references within
% a given article.
